Constraints on the dark matter halo mass of Lyman-alpha emitting galaxies at $z$=3.1 from cosmic variance

### 2013-02-01 paper/ms.tex
updated text

\subsection{The Observational Constraints}
We take as a reference the recently published results of a panoramic survey of LAES at z=3.1 by Yamada et al. 2012. This survey was conducted with the Subaru 8.2m telescope and the Subaru Prime Focus Camera, which has a field of view covering $34\times 27$ arcmin, corresponding to a comoving scale of $46\times35$ Mpc $h^{-1}$ at $z=3.09$. The narrow band filter is centered at $4977$ \AA with a $77$\AA width, corresponding to the redshift range $z=3.062-3.125$ and $41$ Mpc $h^{-1}$ comoving scale for the detection of the Lyman-$\alpha$ line centered at $z=3.09$.

In this paper we only use the results of Yamada et al to constraint our model. This choice is made because their surveys is the largest in area with a set of homogenous conditions that define the LAE sample. Other surveys by XXX an XXX that cover similar regions, but they use different criteria on the equivalent width (EW) cuts to construct the LAE samples. Different cuts in the EW can change the number of LAEs to be included in the catalog. This cuts have an impact on the fainter LAEs which are more abundante than brighter ones. Different definitions of the EW cuts can yield number densities different by a factor of two [REF, I think Yamada has some numbers].

@@

The Bolshoi simulation was performed in a cubic volume of 250 $h^{-1}$ Mpc on a side. It includes dark matter distribution is sampled using $2048^{3}$ particles, which translates into a particle mass of $m_{\rm p}=1.35\times 10^{8}$ $h^{-1}$ M$_{\odot}$.

The cosmological parameters are consistent with a WMAP5 and WMAP7 data with a matter density $\Omega_{\rm m} = 0.27$, cosmological constant $\Omega_{\Lambda}=0.73$, dimensionless Hubble constant $h=0.70$, slope of the power spectrum $n=0.95$ and normalization of the power spectrum $\sigma_{8}=0.82$ [REF].

We have used halo catalogs constructed with a Friend-of-Friends (FOF) algorithm with a linking lenght of 0.17 times the interparticle distance. We have veryfied that the main results we present in this paper also hold if instead we use halo catalogs constructed from a the Bound Density Maxima (BDM) algorithm \citep{KlypinBDM} that are defined to have an density of 200 times the critical density. The minimum halo mass in the models we construct in this Letter correspond to groups of $\sim 75$ particles.

The catalogs we have used here are obtained from a publicly available Multidark database \footnote{{\tt http://www.multidark.org/MultiDark/}} \citep{2011arXiv1109.0003R}.

@@

The model that populates halos with LAEs is based one-to-one correspondence: each halo can only host a single LAE. There are three physical parameters in the model: the halo mass range $M_{\rm min}< M_{\rm halo} < M_{\rm max}$ where LAEs reside and the fraction $f_{\rm occ}$ of such halos that effectively host a LAE. In what follows we will describe by the letter ${\mathcal M}$ a model spanning a set of mock catalogs calculated by fixed values of these three parameters.

We highlight that we do not intent to guess the luminosity of each LAE. We only care about its existence. Physically speaking we are interested in constraining the halo mass above which a LAE can be observed under the conditions defined by Yamada et al.

\subsection{Making Mock Fields}

We create LAEs mock fields from disjoing volumes in the simulation with the same geometry probed by Suprime-CAM and the narrow band filter of $46\times 35\times 41$ $h^{-3}Mpc^{3}$ . There are in total $5\times 7 \times 6=210$ of such sub-volumes in a snapshot of the Bolshoi simulation. We take the $z$ direction to correspond to the redshift direction in the survey. This corresponds to a total area of $880$ arcmin$^{2}$ for each mock field.

Next we group the mock fields in three different ways to construct the LAEs number density distribution The first way we fully take into account the observational constraints and constructs $30$ different mock surveys, each one composed of $12$ mock fields, out of which $7$ correspond to contiguous sub-boxes in the simulation to mimick XXX, $3$ are also contiguous between them but not to the first $7$ fields to mimick XXX and finally $2$ non-contiguous fields that correspond to the XXX data.

From this set of sub-volumes we create groups of we create two different kinds of mock surveys. The first one is constructed to match the observatinal geometry of the observational fields included in this

\subsection{Model Sampling and Selection}
We generate a series of models ${\mathcal M}_{I}$, where $I$ corresponds to set of input values $I=\{M_{\rm min}, M_{\rm max}, f_{\rm occ}\}$. We explore a wide range of values in the $I$ parameter space as follows. $M_{\rm min}$ and $M_{\rm max}$ are allowed to take 15 different values evenly spaced by $0.2$ dex, $M_{\rm min}$ ranges from $\log_{10}M_{\rm min}=10.0$ up to $\log_{10}M_{\rm min}=12.8$, while $M_{\rm max}$ range  from $\log_{10}M_{\rm min}=10.2$ up to $\log_{10}M_{\rm min}=13.0$ shifted by $0.2$ dex. The occupation fraction $f_{\rm occ}$ takes 100 different values from $0.01$ to $1.00$ regularly spaced by $0.01$. In total the number of different sets of input parameters to be explored is $15\times 15\times 100 = 22500$.

For each model we generate 210 mock catalogs corresponding to each of the sub-volumes extracted in the box. For each volume we project along the $z$ direction and calculate its surface number density in units of sources per arcmin$^{2}$. Using the $210$ values for the surface density we perform a Kolmogorov-Smirnov test to compare it against the $12$ observational values and obtain the probability to reject the null hypothesis that the two data sets come from the same distribution.

We also boot-strap these 210 mocks for each model to construct sub-samples of 12 catalogs corresponding to the 12 observed fields. We do this boot-strapping in two different ways. First, by selecting the 12 catalogs in a random fashion without taking into account any spatial information. Second, by selecting sets of catalogs taking into account the spatial correlations present in observations. For instance, 7 out of the 12 catalogs are selected to be spatially contiguous as the SSA22 field, 3 fields correspond to the SXDS fields and 2 remaining catalogs are spatially uncorrelated to resemble the SDF and GOODS-North fields. We do this boot-straping 15 times and keep the median value for the KS test.

>>>>

\subsection{The Observational Constraints}

Our observational reference are the recently published results of a panoramic survey of LAES at z=3.1 by Yamada et al. 2012. This survey was conducted with the Subaru 8.2m telescope and the Subaru Prime Focus Camera, which has a field of view covering $34\times 27$ arcmin, corresponding to a comoving scale of $46\times35$ Mpc $h^{-1}$ at $z=3.09$. The narrow band filter is centered at $4977$ \AA with a $77$\AA width, corresponding to the redshift range $z=3.062-3.125$ and $41$ Mpc $h^{-1}$ comoving scale for the detection of the Lyman-$\alpha$ line centered at $z=3.09$.

The choice to have only one the data from Yamada et al 2012 as reference was  made because their surveys is the largest in area with a set of homogenous conditions that define the LAE sample. Other surveys by XXX an XXX that cover similar regions, but they use different criteria on the equivalent width (EW) cuts to construct the LAE samples. Different cuts in the EW can change the number of LAEs to be included in the catalog. This cuts have an impact on the fainter LAEs which are more abundante than brighter ones. Different definitions of the EW cuts can yield number densities different by a factor of two [REF, I think Yamada has some numbers].

@@

The Bolshoi simulation was performed in a cubic volume of 250 $h^{-1}$ Mpc on a side. It includes dark matter distribution is sampled using $2048^{3}$ particles, which translates into a particle mass of $m_{\rm p}=1.35\times 10^{8}$ $h^{-1}$ M$_{\odot}$.  The cosmological parameters are consistent with a WMAP5 and WMAP7 data with a matter density $\Omega_{\rm m} = 0.27$, cosmological constant $\Omega_{\Lambda}=0.73$, dimensionless Hubble constant $h=0.70$, slope of the power spectrum $n=0.95$ and normalization of the power spectrum $\sigma_{8}=0.82$ [REF].

We use halo catalogs constructed with a Friend-of-Friends (FOF) algorithm with a linking lenght of 0.17 times the interparticle distance. We have veryfied that the main results we present in this paper also hold if instead we use halo catalogs constructed from a the Bound Density Maxima (BDM) algorithm \citep{KlypinBDM} that are defined to have an density of 200 times the critical density. The minimum halo mass in the models we construct in this Letter correspond to groups of $\sim 75$ particles. The catalogs were obtained from the publicly available Multidark database \footnote{{\tt http://www.multidark.org/MultiDark/}} \citep{2011arXiv1109.0003R}.

@@

The model that populates halos with LAEs is based one-to-one correspondence: each halo can only host a single LAE. There are three physical parameters in the model: the halo mass range $M_{\rm min}< M_{\rm halo} < M_{\rm max}$ where LAEs reside and the fraction $f_{\rm occ}$ of such halos that effectively host a LAE. In what follows we will describe by the letter ${\mathcal M}$ a model defined by these three parameters $M_{\rm min}$, $M_{\rm max}$ y $f_{\rm occ}$.

We stress that we do not intent to build a model for the luminosity of each LAE. Physically speaking we are primarely interested in constraining the halo mass above which there are detectable LAEs. under the conditions defined by Yamada et al.

For each mode ${\mathcal M}$ we create mock field from disjoint volumes in the simulation with the same geometry probed by Suprime-CAM and the narrow band filter, namely $46\times 35\times 41$ $h^{-3}$Mpc$^{3}$ where the last dinemsion goes in the redshift direction, corresponding to a total area of $880$ arcmin$^{2}$ for each mock field. There is a total $5\times 7 \times 6=210$ of such sub-volumes in a snapshot of the Bolshoi simulation.

Next we group these $210$ mock fields in three different ways to construct the LAEs number density distribution. The first way we follow the observational setup and constructs $15$ different mock surveys, each one composed of $12$ mock fields, out of which $7$ correspond to contiguous sub-boxes in the simulation to mimick the whole SSA22, $3$ are also contiguous between them but not to the first $7$ fields to mimick the SXDS fields and finally $2$ non-contiguous fields that correspond to the SDF and GOODS-North fields. This will produce $15$ different distributions for the number density for a given model ${\mathrm M}$. The second is similar to the first one. There aare $15$ different mock surveys with $12$ mock fields each, but this time each field corresponds to uncorrelated sub-boxes in the simulation. The third way in only has $1$ mock survey containing all the $210$ mock fields, in this setup there is only one predicted number density distribution for each model ${\mathrm M}$

The advantage of these three ways of sampling is that it allows us to explore the effects of both cosmic variance and the correlation between fields. Comparing the results of the first and second method will help us to quantify the effect of field correlation, while compating the first and the third method will serve us to gauge the impace of cosmic variance.

\subsection{Model Sampling and Selection}

We generate a series of models ${\mathcal M}$ with different input parameters$\{M_{\rm min}, M_{\rm max}, f_{\rm occ}\}$. We explore a wide range of values in the parameter space as follows. $M_{\rm min}$ and $M_{\rm max}$ are allowed to take 30 different values evenly spaced by $0.1$ dex, $M_{\rm min}$ ranges from $\log_{10}M_{\rm min}=10.0$ up to $\log_{10}M_{\rm min}=12.9$, while $M_{\rm max}$ range  from $\log_{10}M_{\rm min}=10.1$ up to $\log_{10}M_{\rm min}=13.0$. The occupation fraction $f_{\rm occ}$ takes 100 different values from $0.01$ to $1.00$ regularly spaced by $0.01$. In total the number of different sets of input parameters to be explored is $30 \times 30\times 100 = 290000$.

For each model we generate the number density distributions in three different ways as described in the previous section. For each volume we project along the $z$ direction and calculate its surface number density in units of sources per arcmin$^{2}$. For each number density distributionwe perform a Kolmogorov-Smirnov test to compare it against the $12$ observational values. From this test we obtain the value $0<P<1$ to reject the null hypothesis that the two data sets come from the same distribution. In this paper we use values of $P<0.1$ to consider that the simulated and observed number densities do not come from the same distributions.
